Gringo 2.0

### 2008-08-04 doc/documentation/paper.tex
additional bitex references, thx to martin, small changes
git-svn-id: https://potassco.svn.sourceforge.net/svnroot/potassco/branches/gringo2@418 3e02f596-3131-0410-8352-e04eeaec37cc

\begin{document}
\maketitle
\tableofcontents
\section{Introduction}
Answer Set Programming (ASP) \cite{baral02a} is a declarative approach for modeling and solving search problems, represented as logic programs.
An ASP system usually consists of a grounder and a solver.
First, a grounder translates a non-ground problem description into a propositional program, which can be processed by the second component of the system, the solver. This is illustrated in figure (\ref{fig:ASP}).
\begin{figure}[h]
\[
\xymatrix{*+[F]\txt{Problem\\representation}\ar[rr]^{Grounder}&&
		*+[F]\txt{Intermediate\\ representation} \ar[rr]^{Solver}&&
		*+[F]\txt{Variable\\ assignment}}
\]
\caption{Architecture of an ASP System}
\label{fig:ASP}
\end{figure}
gringo2.0 is a state of the art grounder and has many features that makes writing logic programs easy.
It was developed by Roland Kaminski and is orientated on the ideas of gringo1.0.
Therefore it supports Bindersplitting, Backjumping and all other features already implemented in gringo1.0.\cite{gringopaper}
The goal of this work is to give an overview of the features supported by gringo2.0. A broad set of potentialities for writing encodings for logic programs exists. Currently, no other grounder supports such a rich featured language.
gringo2.0 supports several output formats and even a human readable format for debugging purposes.
It also has an internal interface to the solver clasp \cite{gekanesc07b}, so the output can directly by parsed to the solver and answer sets can be computed.
Furthermore gringo2.0 can be used with clasp incrementally. Detailed information about incremental solving can be found here \cite{IncrementalASP} 
First, the general restrictions to the input language are stated. Then a brief description of how to use gringo2.0 is given.
In the next chapter the syntax of the supported language is described in detail and a short introduction of how to write a logic program is given. The paper closes with some examples of logic programs and gives an outlook of future work that has to be done in the area of grounding logic programs.
The syntax of the gringo2.0 input language can be found in the appendix \ref{appendix}.

@@

\paragraph{\texttt{--convert}}
If this parameter is used, the grounder expects an already grounded program as input.
For converting an already grounded logic program into an output format supported by gringo2.0 this is the right option. This mode saves a lot of memory but does not ground the program.
\paragraph{\texttt{-l}}
This mode prints smodels conform output. For a definition of the output format see \cite{lparseManual}.
Since some features of gringo2.0 are not natively supported for this output language, new atoms are introduced to create an equivalent program.\newline
This is the current default output format of gringo2.0.
Sum aggregates with negative weights are handled like weight constraint in lparse for this output format. So a transformation as explained in \cite{lparseManual} is used to get rid of the negative weights.
\paragraph{\texttt{-c}}
In this mode, the output is directly parsed to clasp, an ASP solver. For solving logic programs, no intermediate files are produced and the answer sets are printed directly.
With this option enabled, all parameters for clasp are valid and can be declared in the command line. Therefore a double minus ``\texttt{--}'' must be given after all gringo2.0 related parameters are set. After them the clasp parameters follow.
For example:
\begin{verbatim}
 gringo -c --const n=5 -- --supp-models 3
\end{verbatim}
Furthermore all restriction on smodels output hold as clasps internal interface currently reads smodels format.
\paragraph{\texttt{-p}}
Parameter ``\texttt{-p}'' invokes the plain output method, so the output is printed in human readable format, which is also the input format for gringo2.0. The output is variable free and some simplifications on the program were made. To convert the plain format into a format readable by a solver use the \texttt{--convert} option.
\paragraph{\texttt{-g [1-7]}}
This parameter invokes the new output format for the intermediate languages ASPils \cite{atleastthisreferenceishouldgive}. The optional number can range from $1$ to $7$ and refers to one of the following normal forms (Simple=1, SModels=2, DisjunctiveLogicProgram=3, CModels=4, CModelsExtended=5, DLV=6, Conglomeration=7). If no normal form is specified, conglomeration(7)  is assumed.

@@

\subsection{Meta-Syntax}
The ``\#'' sign in front of all meta expressions is optional but is it strongly recommend to use it.
It separates the meta directives from the logic program more clearly and can prevent severe mistakes,
 confusing those constructs with atoms as shown in the example.
\begin{verbatim}
 hide.
 hidden :- hide.
\end{verbatim}
This program hides all predicates from the output of the ASP solver instead of deriving the facts \texttt{hide} and \texttt{hidden}.

@@

#const queens = 200.
#const diags  = queens * 2 - 1.

d(1..queens).
diag(1..diags).

>>>>

\begin{document}
\maketitle
\tableofcontents
\section{Introduction}
Answer Set Programming (ASP) \cite{baral02a} is a declarative approach for modeling and solving search problems, represented as logic programs.
An ASP system usually consists of a grounder and a solver.
First, a grounder translates a non-ground problem description into a propositional program, which can be processed by the second component of the system, the solver. This is illustrated in figure (\ref{fig:ASP}).
\begin{figure}[h]
\[
\xymatrix{*+[F]\txt{Problem\\representation}\ar[rr]^{Grounder}&&
		*+[F]\txt{Intermediate\\ representation} \ar[rr]^{Solver}&&
		*+[F]\txt{Variable\\ assignment}}
\]
\caption{Architecture of an ASP System}
\label{fig:ASP}
\end{figure}
gringo2.0 is a state of the art grounder and has many features that makes writing logic programs easy.
It was developed by Roland Kaminski and is orientated on the ideas of gringo1.0.
Therefore it supports Bindersplitting, Backjumping and all other features already implemented in gringo1.0.\cite{gringopaper}
The goal of this work is to give an overview of the features supported by gringo2.0. A broad set of potentialities for writing encodings for logic programs exists. Currently, no other grounder supports such a rich featured language.
gringo2.0 supports several output formats and even a human readable format for debugging purposes.
It also has an internal interface to the solver clasp \cite{gekanesc07b}, so the output can directly by parsed to the solver and answer sets can be computed.
Furthermore gringo2.0 can be used with clasp incrementally. Detailed information about incremental solving can be found here \cite{gekaosscth08a} 
First, the general restrictions to the input language are stated. Then a brief description of how to use gringo2.0 is given.
In the next chapter the syntax of the supported language is described in detail and a short introduction of how to write a logic program is given. The paper closes with some examples of logic programs and gives an outlook of future work that has to be done in the area of grounding logic programs.
The syntax of the gringo2.0 input language can be found in the appendix \ref{appendix}.

@@

\paragraph{\texttt{--convert}}
If this parameter is used, the grounder expects an already grounded program as input.
For converting an already grounded logic program into an output format supported by gringo2.0 this is the right option. This mode saves a lot of memory but does not ground the program.
\paragraph{\texttt{-l}}
This mode prints smodels conform output. For a definition of the output format see \cite{lparseManual}.
Since some features of gringo2.0 are not natively supported for this output language, new atoms are introduced to create an equivalent program.\newline
This is the current default output format of gringo2.0.
Sum aggregates with negative weights are handled like weight constraint in lparse for this output format. So a transformation as explained in \cite{lparseManual} is used to get rid of the negative weights.
\paragraph{\texttt{-c}}
In this mode, the output is directly parsed to clasp, an ASP solver. For solving logic programs, no intermediate files are produced and the answer sets are printed directly.
With this option enabled, all parameters for clasp are valid and can be declared in the command line. Therefore a double minus ``\texttt{--}'' must be given after all gringo2.0 related parameters are set. After them the clasp parameters follow.
For example:
\begin{verbatim}
 gringo -c --const n=5 -- --supp-models 3
\end{verbatim}
Furthermore all restriction on smodels output hold as clasps internal interface currently reads smodels format.
\paragraph{\texttt{-p}}
Parameter ``\texttt{-p}'' invokes the plain output method, so the output is printed in human readable format, which is also the input format for gringo2.0. The output is variable free and some simplifications on the program were made. To convert the plain format into a format readable by a solver use the \texttt{--convert} option.
\paragraph{\texttt{-g [1-7]}}
This parameter invokes the new output format for the intermediate languages ASPils \cite{gejaosscth08b}. The optional number can range from $1$ to $7$ and refers to one of the following normal forms (Simple=1, SModels=2, DisjunctiveLogicProgram=3, CModels=4, CModelsExtended=5, DLV=6, Conglomeration=7). If no normal form is specified, conglomeration(7)  is assumed.

@@

\subsection{Meta-Syntax}
This section is about constructs that do not originally belong to the logic program but can be used to give additional information for example to the solver.
The ``\#'' sign in front of all meta expressions is optional but is it strongly recommend to use it.
It separates the meta directives from the logic program more clearly and can prevent severe mistakes,
 confusing those constructs with atoms as shown in the example.
\begin{verbatim}
 hide.
 hidden :- hide.
\end{verbatim}
This program hides all predicates from the output of the ASP solver instead of deriving the facts \texttt{hide} and \texttt{hidden}.

@@

% this can be set via the command line
#const queens = 200.

#const diags  = queens * 2 - 1.
d(1..queens).
diag(1..diags).
